---
title: "Base Plot in R"
author: "Gregoire Michoud"
date: "02/01/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Base Plot in R

One of the most powerful functions of R is it's ability to produce a wide range of graphics to quickly and easily visualise data. Plots can be replicated, modified and even publishable with just a handful of commands.

For this, we are going to use the `airquality` dataset, which measure the daily air quality in New York from May to September 1973, with 6 variables :

* Ozone: Mean ozone in parts per billion from 1300 to 1500 hours at Roosevelt Island
* Solar.R: Solar radiation in Langleys in the frequency band 4000–7700 Angstroms from 0800 to 1200 hours at Central Park
* Wind: Average wind speed in miles per hour at 0700 and 1000 hours at LaGuardia Airport
* Temp: Maximum daily temperature in degrees Fahrenheit at La Guardia Airport.
* Month
* Day

```{r}
data(airquality)
head(airquality)
```


As you can see all variables are numerical, so we will create a factorial variable for plotting purposes later on

```{r}
airquality$Solar.R_cat<-cut(airquality$Solar.R, c(0,50,100,150,200,250,300,350))
airquality$Solar.R_cat
```

### Histograms

A very simple histogram to find the frequency of the temperature

```{r}
hist(airquality$Temp)
```

A more complex one with the title, x label and we changed the number of cells to 20

```{r}
hist(airquality$Temp, breaks = 20, xlab = "Temperature (degrees F)",
     main = "Histogram of Temperature in New from May to September 1973",
     ylab = "Frequency of Temperature",
     ylim = c(0,20),
     xlim = c(50,100)
)
```

### Scatterplot

A simple scatter plot to compare the ozone concentration and the wind

```{r}
plot(x = airquality$Ozone, y = airquality$Wind)
```

It seems to be negatively correlated, does the solar radiation play a role ?

```{r}
plot(x = airquality$Ozone,
     y = airquality$Wind,
     col = airquality$Solar.R_cat,
     pch = 16,
     xlab = "Ozone in parts per billion", 
     ylab = "Average wind speed (mph)")
legend(title = "Solar radiation (lang)", x = 120, y = 20, legend = levels(airquality$Solar.R_cat), col = c(1:7), pch = 16)
```

### Boxplot

Another type of plot

```{r}
boxplot(airquality$Wind ~ airquality$Solar.R_cat,
        ylab = "Average wind speed (mph)",
        xlab = "Solar radiation (lang)")
```
