---
title: "Colors in R"
author: "Tamara Huete-Stauffer"
date: "12/15/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RColorBrewer)
```

## Colors in R

Here you can find the available colors in R
<http://research.stowers.org/mcm/efg/R/Color/Chart/>

And here is a web page that helps you decide which palette to use
<http://colorbrewer2.org/>

#### 1. Accessing individual colors
```{r naming colors}
## 3 ways of accessing individual colors: 
### by name   = "aquamarine4" 
### by number = colors()[12]
### by hexadecimal code = "#458B74"

plot(x=1:10, y=1:10, col="aquamarine4", pch=15,cex=1.2) ## name
points(x=1:10,y=2:11, col=colors()[12], pch=16,cex=1.2) ## number
points(x=1:10, y=3:12, col="#458B74",pch=17,cex=1.2) ## hex code
legend(x=8, y=4, legend=c("aquamarine4","colors()[12]","#458B74"),pch=c(15,16,17), col=c(colors()[12],"aquamarine4","#458B74"))


```

#### 2. Using RColorBrewer palettes
```{r display}
## These are the available built-in palettes
par(cex=0.7) ## cex regulates the size of the font
display.brewer.all()

## Display color blind friendly palettes
display.brewer.all(colorblindFriendly=TRUE)

## View a single RColorBrewer palette by specifying its name and number of colors you need
par(mfrow=c(1,3))
display.brewer.pal(n=9,name = 'RdBu')
display.brewer.pal(n=11,name = 'Spectral')
display.brewer.pal(n=5,name = 'Spectral')

```

Select the number of colors you need depending on the number of cases in your plot
```{r bars}
## Select the number of colors you need.
## If you select less colors than the number of x-y points, they will be repeated
par(mfrow=c(1,2)) ### make a 2 by 2 window to plot 4 plots at the same time
barplot(c(1,2,5,7,6,7), col=brewer.pal(n = 6, name = "Spectral"),main="Spectral, n = 6")
barplot(c(1,2,5,7,6,7), col=brewer.pal(n = 3, name = "Spectral"), main="Spectral, n = 3")
brewer.pal(n=9,name ="Spectral")

```

Use colorRampPalette() to extend the maximum number of breaks in the built-in palettes
```{r colorRampPalette}

## the palettes have a maximum number of breaks, but you can extend them with colorRampPalette()
Pal100 <- colorRampPalette(brewer.pal(9,"Spectral"))(100) ### 100 colors of the Spectral palette
Pal50 <- colorRampPalette(brewer.pal(9,"Spectral"))(50) ### 50 colors of the Spectral palette

par(mfrow=c(1,2))
plot(x=1:100,y=1:100,col=Pal100,cex=2,pch=19, main="Spectral, n=100") ## 100 Spectral palette
plot(x=1:100,y=1:100,col=colorRampPalette(brewer.pal(9,"Blues"))(100),cex=2,pch=19,main="Blues, n=100") ## 100 Blues palette


```

Or make your own palettes!!
```{r customize palette}
## or you can make your own
myPal <- colorRampPalette(c("blue","cyan","orange","firebrick1"))

par(mfrow=c(1,2))
plot(x=1:10,y=1:10,col=myPal(10),cex=2,pch=19, main="myPal, n=10")
plot(x=1:100,y=1:100,col=myPal(100),cex=2,pch=19,main="myPal, n=100")
```

